---
title: "Showcase - Election App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Showcase - Election App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
In this article we demonstrate how easy it is to connect Graphs with a `leafdown` map.
TODO add image

## Data
As described in the Introduction, we need two types of data:

**SpatialPolygonsDataFrames**, the shapes of the US-States and Counties, taken from the `raster` package.

**Election Results and Census Data**, the data we want to display on the map, taken from the example data sets that come with the `leafdown` package.
(The original data comes from [Deleetdk](https://github.com/Deleetdk/USA.county.data). For more information about the data, please see TODO change `?us16_1` or `?us16_2` respectively)

## The Map itself
The structure of the map is pretty similar to the map from the Introduction article.
Here we show the results of the US Presidential Election from 2016.

## Connecting Graphs to our Map
In this section we want to demonstrate how simple it is to connect graphs or similar UI-elements with the map.

We create two graphs that give more insight on the currently selected shapes. </br>
- As the map only shows the winner, we create a barchart to show the percentages from every party. </br>
- Additionally we add a graph showing the racial makeup.

The changes in the UI are straight forward:
```{r, eval = FALSE}
column(
  width = 5,
  # box for racial makeup graph
  bs4Card(
    width = 12,
    closable = F,
    collapsible = F,
    title = "Racial makeup in percentages",
    echarts4rOutput("socio")
  ),
  # box for party percent graph
  bs4Card(
    width = 12,
    closable = F,
    collapsible = F,
    title = "Votes in percent",
    echarts4rOutput("party")
  )
)
```

To connect the graphs with the map, we can use the `$curr_sel_data()` attribute.
This attribute is a `reactiveValue` which allows us to update the graphs whenever the user selects a shape on the map or drills a level up or down.

In the server we get the data using `df <- my_leafdown$curr_sel_data()`. 

Creating the rest of the graph is again straight forward.

```{r, eval = FALSE}
output$party <- renderEcharts4r({
  # get the currently selected data from the map
  df <- my_leafdown$curr_sel_data()
  
  # check whether any shape is selected, show general election-result if nothing is selected
  if(dim(df)[1] > 0){
    if(my_leafdown$curr_map_level == 1) {
      df <- df[, c("state_abbr", "Democrats2016", "Republicans2016", "Libertarians2016", "Green2016")]
      df <- df %>% pivot_longer(2:5, "party") %>% group_by(party)
    } else {
      df <- df[, c("County", "Democrats2016", "Republicans2016", "Libertarians2016", "Green2016")]
      df <- df %>% pivot_longer(2:5, "party") %>% group_by(party)
      df$value <- df$value / 100
      names(df)[1] <- "state_abbr"
    }
  } else {
    # show general election-result as no state is selected
    df <- data.frame(
      party = c("Democrats2016", "Republicans2016", "Libertarians2016", "Green2016"),
      state_abbr = "USA",
      value = c(0.153, 0.634, 0.134, 0.059)) %>% group_by(party)
  }
  # create the graph
  df %>%
    e_charts(state_abbr, stack="grp") %>%
    e_bar(value) %>%
    e_y_axis(formatter = e_axis_formatter("percent", digits = 2)) %>%
    e_tooltip(trigger = "axis",axisPointer = list(type = "shadow")) %>%
    e_legend(right = 10,top = 10) %>%
    e_color(c("#232066", "#E91D0E", "#f3b300", "#006900"))%>%
    e_tooltip(formatter = e_tooltip_item_formatter("percent", digits = 2))
})
```

# TODO add map demo
